---
title: "lab 1 fr"
author: "Maya Szafraniec"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# attach packages
```{r}
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```


# read in data frame about the fisheries 

```{r}
us_landings <-read_csv(here("data","noaa_fisheries.csv"))

```

Now, make a tidy version of the data frame

parce number allows us to seperate the number from things when there is not somethign that divides the number from the unit, just pull out the number without making a new colum for the thing that you are removing (the dollar sign in this example)

```{r}
landings_tidy <-us_landings %>% 
  clean_names()%>% 
  mutate(state = str_to_lower(state),
         afs_name = str_to_lower(afs_name)) %>% 
  mutate(dollars_num = parse_number(dollars_usd))
```


Lets get some information for salmon

string remove is for taking out the things that are not helpful for the current analysis 

filter part search data for when at least part of the pattern is in the data

Seperate, when you get to these, split things into these two columns

```{r}
salmon_landings <-landings_tidy %>%
  mutate(afs_clean = str_remove(afs_name, patter = "aggregate")) %>% 
  filter(str_detect(afs_clean, pattern = "salmon")) %>% 
  separate(afs_clean, into = c("group","species", sep = ","))
  
```
Find some grouped summary data:

group by and summarize to make groups and then tell them how to seperate it

find annual total US landing and dollar value (summing across all states) for each type fo salmon 

```{r}
salmon_summary<-salmon_landings %>% 
  group_by(year,species) %>% 
  summarize(
    tot_landings = sum(landings_pounds),
    tot_value = sum(dollars_num)
  )

```

Make a graph!

```{r}
salmon_landings_graph <- ggplot(data = salmon_summary, aes(x=year,y=tot_landings))+
  geom_line(aes(color=species))+
  theme_minimal()

salmon_landings_graph


```

Now we will export our graph with ggsave (put it back in the figures folder in the repo, tell it to put it here and name it this)

```{r}
ggsave(plot = salmon_landings_graph,
       here("figures","us_salmon_ms.png"),
       height =5,
       width =8)
```

Make a kable table they are nice
top 5 lines

```{r}
salmon_first_5<-salmon_summary %>% 
  head(5)
```

Striped makes every other line dark
full width makes it not fill up the whole html doctumet

```{r}
kable(salmon_first_5) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```


